Arrows for Parallel Computations

### 2017-12-17 paper/content/motivation.tex
minor
minor fix in motivation.tex

\section{Introduction}
\label{sec:introduction}
% \olcomment{todo, reuse 5.5, and more}
%
% \mbcomment{
% Haskell is Spielwiese für Parallelität; verschiedene Ansätze (Par, Multicore, Eden); Orthogonale Ansätze; Verwenden höchstens eine Monade - manchmal auch nur intern; Wir wollen Parallelität mit Arrows abbilden, was noch niemand gemacht hat;
% Statt einer eigenen Implementierung definieren wir ein "shallow embedded DSL" (ACHTUNG, ist das der richtige Name? effektiv API);
% Umsetzung mit verschiedenen parallelen Haskells; We tame the zoo of parallel Haskells und vergewissern uns dass es nicht viel Overhead bringt
% }
%
% blablabla arrows, parallel, haskell.
%
% \olcomment{An attempt ---->}
%%%%%

@@

%In this paper we introduce a notion of parallel
%computations using Arrows.
A key novelty in this paper is to use Arrows to represent parallel computations. They seem a natural fit as they are a generalization of the function arrow (|->|) and serve as general interface to computations. Section \ref{sec:arrows} gives a short introduction to Arrows.

%Our Arrows-based interface is a high-level one.
We provide an Arrows-based type class and implementations for the three above mentioned parallel Haskells.
Instead of 
introducing a new low-level parallel backend in order to implement our
Arrows-based interface, we define a shallow-embedded DSL for Arrows. This DSL
is defined as a common interface with varying implementations in
the existing parallel Haskells. Thus, we not only define a parallel programming interface in a
novel manner -- we tame the zoo of parallel Haskells. We provide a
common, very low-penalty programming interface that allows to switch
the parallel backends at will. The induced penalty was max. 8.3 \% in our measurements \mbcomment{, and typically under X \%}. Further backends based on HdpH or a Frege implementation (on the Java Virtual Machine) are viable, too.

\paragraph{Contributions}
%
%\olcomment{HIT HERE REALLY STRONG}
%
%\subsection{Impact of parallel Arrows}
%\olcomment{move this to Contributions in the front or something}
We propose an Arrow-based encoding for parallelism based on a new Arrow combinator |parEvalN :: [arr a b] -> arr [a] [b]| in Section \ref{sec:parallel-arrows}, which converts a list of Arrows into a new parallel Arrow. A parallel Arrow is still an Arrow, hence the resulting parallel Arrow can still be used in the same way as a potential sequential version.

In this paper evaluate the expressive power of our Arrow formalism in the context of parallel programming.

@@

\mbcomment{Überleitung auf Related Work hier machen}

\mbcomment{Mention Future work here?}

\mbcomment{Conclusion?}

>>>>

\section{Introduction}
\label{sec:introduction}

@@

A key novelty in this paper is to use Arrows to represent parallel computations. They seem a natural fit as they can be thought of as a more general function arrow (|->|) and serve as general interface to computations while not being as restrictive as Monads \citep{HughesArrows}. Section \ref{sec:arrows} gives a short introduction to Arrows.

We provide an Arrows-based type class and implementations for the three above mentioned parallel Haskells.
Instead of 
introducing a new low-level parallel backend in order to implement our
Arrows-based interface, we define a shallow-embedded DSL for Arrows. This DSL
is defined as a common interface with varying implementations in
the existing parallel Haskells. Thus, we not only define a parallel programming interface in a
novel manner -- we tame the zoo of parallel Haskells. We provide a
common, very low-penalty programming interface that allows to switch
the parallel backends at will. The induced penalty was max. 8.3 \% in our measurements \mbcomment{, and typically under X \%}. Further backends based on HdpH or a Frege implementation (on the Java Virtual Machine) are viable, too.

\paragraph{Contributions}
%
%\olcomment{HIT HERE REALLY STRONG}
%
%\subsection{Impact of parallel Arrows}
%\olcomment{move this to Contributions in the front or something}
We propose an Arrow-based encoding for parallelism based on a new Arrow combinator |parEvalN :: [arr a b] -> arr [a] [b]|. A parallel Arrow is still an Arrow, hence the resulting parallel Arrow can still be used in the same way as a potential sequential version. In this paper we evaluate the expressive power of such a formalism in the context of parallel programming.

We structure this paper as follows:

@@

PArrows are open source and available from \url{https://github.com/s4ke/Parrows}.
